---
title: "02_Data-analysis (Mult)"
format: html
editor: visual
---

## Library

```{r}
#| message: False

library(lme4)
library(lmerTest)
library(tidyverse)
library(performance)
library(emmeans)
library(ggeffects)
library(ggplot2)
```

### Import

```{r}

participant_info <- read_csv("participants_demo.csv", show_col_types = FALSE) |>
  select(-(id:testing_date)) |>
  mutate(mult_fluency_group = factor(mult_fluency_group, 
                                   levels = c("Low", "High")))

### ------------------------------------------------

folder_path <- "RAW-DATA_mult_behavioral"

all_data <- list.files(path = folder_path, pattern = "\\.csv$", full.names = TRUE) %>% 
  map_dfr(read_csv, show_col_types = FALSE)

df <- all_data |>
    select(participant, blocks_file,  mult_stim,
    type,
    ProblemSize,
    InterferenceLH,
    InterferenceLevel,
    mult_type,
    Mult_Resp.corr,
    Mult_Resp.rt,
    Strategy_RESP.keys
  ) |>
  filter(!is.na(type)) |>
  rename(
    block_file = blocks_file,           
    problem = mult_stim,            
    problem_size = ProblemSize,             
    interference = InterferenceLH, 
    interference_score = InterferenceLevel,
    stim_type = mult_type,
    rt = Mult_Resp.rt,
    correctness = Mult_Resp.corr,
    strategy_type = Strategy_RESP.keys
  )|>
      mutate(correctness = case_when(
    is.na(rt) ~ 0,                  # if RT is NA, keep as 0
    correctness == 1 ~ 0,           # flip 1 to 0
    correctness == 0 ~ 1            # flip 0 to 1
  ))|>
    mutate(correctness = factor(
    correctness, 
    levels = c(0, 1), 
    labels = c("Incorrect", "Correct")
  ))|>
  mutate(
        type = factor(type, levels = c("small-sized_low-interference", 
                                       "large-sized_low-interference", 
                                   "small-sized_high-interference", 
                                   "large-sized_high-interference")),
    problem_size = factor(problem_size, 
                          levels = c("small-sized", "large-sized")),
    interference = factor(interference, 
                            levels = c("low-interference", 
                                       "high-interference")),
    stim_type = factor(stim_type, 
                       levels = c("CorrectRESP", "IncorrectRESP")),
    strategy_type = factor(strategy_type, 
                         levels = c('a', 'd'), 
                         labels = c("Procedure", "Retrieval"))
  )
```

### Multilevel analysis

```{r}

# reverse coding

df_analysis_acc <- participant_info |>
  inner_join(df, by = c('participant'))

df_analysis <- participant_info |>
  inner_join(df, by = c('participant'))|>
  filter(correctness == "Incorrect") |>  
  mutate(log_rt = log(rt)) 
```

## GLMER (Accuracy)

```{r}

max_correct <- glmer(
  correctness ~ problem_size * interference * mult_fluency_group +
    (1 + problem_size + interference | participant) +
    (1 | problem),
  data = df_analysis_acc,
  family = binomial,
  control = glmerControl(optimizer = "bobyqa")
)

summary(max_correct)
```

### Model Performance

```{r}
# Check random effects structure
performance::model_performance(max_correct)

# See if singular fit
lme4::isSingular(max_correct)

```

### Null model

```{r}
max_correct_null <- glmer(correctness ~ problem_size * interference + 
                            (1 + problem_size + interference | participant) + (1 | problem),
                          data = df_analysis_acc, family = binomial)

anova(max_correct_null, max_correct, test = "Chisq")

```

### log-odds and odds ratios

```{r}

exp(fixef(max_correct)) # odds ratios
exp(confint(max_correct, method = "Wald")) # confidence intervals
exp(cbind(OR = fixef(max_correct), confint(max_correct, method = "Wald")))

```

## GLMER (Retrievals vs. Procedural)

Releveling

```{r}

# Procedure = 0, Retrieval = 1
table(df_analysis_acc$strategy_type)

# Cross-tabulation by each predictor
table(df_analysis_acc$problem_size, df_analysis_acc$strategy_type)
table(df_analysis_acc$interference, df_analysis_acc$strategy_type)
table(df_analysis_acc$mult_fluency_group, df_analysis_acc$strategy_type)
```

Recoding for a stable model (instability due to higher count for retrievals)

```{r}
df_analysis_acc2 <- df_analysis_acc %>%
  mutate(strategy_binary = ifelse(strategy_type == "Procedure", 1, 0))

table(df_analysis_acc2$strategy_binary)

```

### Max Model

```{r}

strategy_model <- glmer(
  strategy_binary ~ problem_size * interference * mult_fluency_group +
    (1 + problem_size + interference | participant) +
    (1 | problem),
  data = df_analysis_acc2,
  family = binomial,
  control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))
)


summary(strategy_model)
```

### log-odds and odds ratios

```{r}

exp(cbind(
  OR = fixef(strategy_model),
  confint(strategy_model, method = "Wald")
))

```

### Plots

```{r}
# Correct the rate: 1 = retrieval, 0 = procedure
summary_df <- df_analysis_acc2 %>%
  filter(!is.na(strategy_binary)) %>%
  group_by(mult_fluency_group, problem_size, interference) %>%
  summarise(
    n = n(),
    retrieval_rate = mean(strategy_binary == 0),  # TRUE for retrievals
    se = sd(strategy_binary == 0) / sqrt(n),
    retrieval_percent = retrieval_rate * 100,
    ymin = retrieval_percent - 1.96 * se * 100,
    ymax = retrieval_percent + 1.96 * se * 100,
    .groups = "drop"
  )


plot_strategy <- ggplot(summary_df, aes(x = problem_size, y = retrieval_percent, color = interference, group = interference)) +
  geom_point(size = 3) +
  geom_line(linewidth = 1.2) +
  geom_errorbar(aes(ymin = ymin, ymax = ymax), width = 0.15) +
  facet_wrap(~ mult_fluency_group) +
  labs(
    title = "Observed Retrieval Strategy Use",
    x = "Problem Size",
    y = "Percentage of Retrievals",
    color = "Interference"
  ) +
  scale_color_manual(values = c("firebrick", "steelblue")) +
  theme_minimal(base_size = 14) +
  theme(
    legend.position = "bottom",
    panel.grid.minor = element_blank()
  )

# Save as PNG
ggsave(
  filename = "retrieval_strategy_plot.png",
  plot = plot_strategy,
  path = "figures",        
  width = 8,
  height = 5,
  dpi = 300
)




```

## LMM (RT)

```{r}

df_rt <- df_analysis_acc2 %>%
  filter(correctness == "Correct") %>%
  mutate(
    log_rt = log(rt)
  )

hist(df_rt$log_rt, breaks = 40)

rt_model <- lmer(
  log_rt ~ problem_size * interference * mult_fluency_group +
    (1 + problem_size * interference | participant) +
    (1 | problem),
  data = df_rt
)

summary(rt_model)


m1 <- lmer(log_rt ~ problem_size * interference * mult_fluency_group + 
           (1|participant) + (1|problem),
           data = df_analysis)

# Random intercepts and random slopes for within-subject factors
m2 <- lmer(log_rt ~ problem_size * interference * mult_fluency_group + 
           (1 + problem_size * interference|participant) + (mult_fluency_group|problem),
           data = df_analysis)

summary(m2)

# Compare models
anova(m1, m2)
```

### emmeans

```{r}
# compare problem size × interference by fluency
emm_options(pbkrtest.limit = 5000)

emm <- emmeans(rt_model, ~ problem_size * interference * mult_fluency_group)
pairs(emm, by = "mult_fluency_group")  
summary(emm, type = "response")  # gives estimates on original RT scale


```

means for each conditions

```{r}

# Get estimated marginal means for each condition (on original RT scale)
emm_summary <- emmeans(rt_model, ~ problem_size * interference * mult_fluency_group)

# Back-transform from log RT to RT in milliseconds
emm_summary_resp <- summary(emm_summary, type = "response")

emm_df <- as.data.frame(emm_summary_resp)

# Print the table
print(emm_summary_resp)

```

### Plots for interactions

```{r}

# Use back-transformed predictions
emm_summary_resp <- summary(emmeans(rt_model, ~ problem_size * interference * mult_fluency_group), type = "response")

# Convert to ms
emm_summary_resp <- emm_summary_resp |>
  mutate(rt_ms = exp(emmean) * 1000,
         lower = exp(lower.CL) * 1000,
         upper = exp(upper.CL) * 1000)

# Plot
rt_emmeans_plot <-ggplot(emm_summary_resp, aes(x = problem_size, y = rt_ms, color = interference, group = interference)) +
  geom_point(size = 3) +
  geom_line(linewidth = 1.2) +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.15) +
  facet_wrap(~ mult_fluency_group) +
  labs(
    title = "Predicted Reaction Time by Problem Size, Interference, and Fluency Group",
    x = "Problem Size",
    y = "Estimated RT (ms)",
    color = "Interference"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    legend.position = "bottom",
    panel.grid.minor = element_blank()
  )

ggsave(
  filename = "figures/rt_emmeans_plot.png",      
  plot = rt_emmeans_plot,               
  width = 8,                             
  height = 5,                            
  dpi = 300                              
)

```

Summary

```{r}
summary(m1)
summary(m2)
```

### posthoc

```{r}
library(emmeans)
emm_options(pbkrtest.limit = 4796)

emm <- emmeans(m2, ~ problem_size | interference | mult_fluency_group)
pairs <- pairs(emm)
pairs_adjusted <- pairs(emm, adjust = "tukey")

summary(emm)
pairs(emm)
pairs_adjusted

```

#### contrast

```{r}

contrasts <- emmeans(m2, pairwise ~ problem_size | interference | mult_fluency_group)

int_contrast <- contrast(emm, interaction = c("pairwise", "pairwise"), 
                        by = "mult_fluency_group")
summary(int_contrast, adjust = "tukey")

```

### plot

```{r}

emm_df <- as.data.frame(emm)
  

interaction_plot <- ggplot(emm_df, aes(x = problem_size, y = emmean, 
                   color = mult_fluency_group, 
                   linetype = interference)) +
  geom_line(aes(group = interaction(mult_fluency_group, interference)), 
            size = 1) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = emmean - SE, 
                    ymax = emmean + SE), 
                width = 0.2) +
  scale_color_brewer(palette = "Set2", name = "Multiplication Fluency") +
  scale_linetype_discrete(name = "Interference") +
  labs(x = "Problem Size",
       y = "Log Reaction Time") +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    panel.grid.minor = element_blank(),
    axis.text = element_text(size = 11),
    axis.title = element_text(size = 12, face = "bold")
  )

ggsave("interaction_plot.png", plot = interaction_plot, width = 10, height = 6, dpi = 600)
```
